Constructing the transition graphs of interacting hysterons

### 2022-06-15 Paper_draft.tex
Corrected several instances of l i/o L
Miscellaneous small corrections after going through draft

The system's response to a field $U$ is thus obtained by comparing the field to the switching fields $\{U_i^+\}, \{U_i^-\}$ of the hysterons individually, flipping hysterons that are unstable, and repeating this process until there are no unstable hysterons. The response is ill-defined if multiple hysterons are unstable at the same time, or if the system becomes caught in an infinite self-loop.\\
The possible responses of a collection of $N$ hysterons to driving are captured by a state transition graph, which consists of all stable states $S$ and all transitions between these states. For a system of $N$ hysterons, there are $2^N$ possible states. In particular there are two saturated states, where either all $s_i=0$ or all $s_i=1$. Each state $S$ has a single up and down transition, except for the two saturated states, which only have an up and a down transition respectively. Therefore, a transition graph for a collection of $N$ hysterons consists of a maximum of $2^N$ states and $2(2^N-1)$ transitions.

@@

The region in parameter space associated with each graph is bounded by a set of linear inequalities that the switching fields must obey, which we refer to as the design inequalities. We generate the design inequalities systematically by combining the conditions on each individual transition (see SI). If the solution space of a graph's design inequalities is not empty, then the graph is realizable. Thus, one can find the possible graphs for $N$ hysterons by first generating the candidate graphs, then checking whether each graph's design inequalities have a solution.\\

@@

It often occurs that one or more of the possible states for $N$ hysterons cannot be reached from the ground state. As is customary, we refer to states that cannot be reached as garden-of-Eden states. Because the transitions from garden-of-Eden states are irrelevant when one starts from the ground state, as is generally the case in our experiments, we may choose to ignore garden-of-Eden states. As a consequence, when two graphs only differ from one another by transitions that occur from garden-of-Eden states, the graphs become duplicates. Therefore, when garden-of-Eden states are ignored, the number of distinct graphs can turn out lower than the number of possible combinations of transitions. For $N=2$, ignoring garden-of-Eden states  reduces the number of graphs from 1458 to 752. 

@@

The idea of base graphs allows us to improve our method for generating candidate graphs. Namely, by first constructing all base graphs, and only then evaluating all possible avalanches for each base graph individually, we exclude the subgraphs of the type shown in Fig. \ref{fig:impossiblegraphs}b from the outset. Because the number of graphs for which the design inequalities need to be explicitly generated is then reduced, the base graph method has an improved efficiency compared to the naive approach. \\

@@

In order to discuss scrambling more explicitly, we here define the main loop as the series of up and down transitions connecting the two saturated states. In the Preisach model, the ordering of the main loop fixes the transitions throughout the transition graph, so that each main loop corresponds to a unique Preisach graph. The number of main loops is the number of permuatations times the number of permutations of the down boundary, which is $N!^2$. However, we fix the up boundary when accounting for exchange symmetry, thus dividing the number of main loops by $N!$.  Therefore, when accounting for exchange symmetry, there are only $N!$ main loops and therefore $N!$ Preisach graphs. In a scrambled graph, one or more of the transitions in the graph deviates from the main loop, so that each main loop corresponds to a number of base graphs (Fig.\ref{fig:basegraphs_n3}).\\

@@

Let a base graph contain the up transition $S^{(0)}\uparrow S^{(1)}$. The state $S^{(1)}$ has two possible transitions: an up transition $S^{(1)}\uparrow S^{(2+)}$, and a down transition  $S^{(1)}\downarrow S^{2-}$.  The transition $S^{(0)}\uparrow S^{(1)}$ can thus be extended to a $l=2$ avalanche in two ways: either by the ferromagnetic avalanche $S^{(0)}\uparrow S^{(1)}\uparrow S^{(2+)}$, or by the antiferromagnetic avalanche $S^{(0)}\uparrow S^{(1)}\downarrow S^{(2-)}$ (Fig. \ref{fig:avalanches}a). Similarly, a down transition $S^{(0)}\downarrow S^{(1)}$ can be extended to the antiferromagnetic avalanche $S^{(0)}\downarrow S^{(1)}\uparrow S^{(2+)}$, or to the ferromagnetic avalanche  $S^{(0)}\downarrow S^{(1)}\downarrow S^{(2-)}$ (Fig. \ref{fig:avalanches}b). Therefore, for each transition $S^{(0)}\rightarrow S^{(1)}$, there is one possible ferromagnetic and one antiferromagnetic avalanche when $S^{(1)}$ is not a saturated state. In the case where $S^{(1)}$ is saturated, only the antiferromagnetic option exists.\\
 We see that every transition $S^{(0)}\rightarrow S^{(1)}$ in the base graph can be extended to a $l=2$ transition  $S^{(0)}\rightarrow S^{(1)}\rightarrow S^{(2)}$ in at most two ways, by choosing either $S^{(2)} = S^{(2+)}$ or $S^{(2)} = S^{(2-)}$ . We must check for each of these transitions that $S^{(2)} \neq S^{(0)}$, because an avalanche cannot contain any cycles. By then evaluating all combinations of valid avalanches, we obtain a list of candidate graphs for each base graph. An example for one of the $N=2$ base graphs is shown in Fig. \ref{fig:N2l2}. For the example shown, the base graph has four possible $l=2$ avalanches all associated with a different transition ($00\uparrow 01\uparrow 11$, $01\uparrow11\downarrow10$, $11\downarrow10\downarrow00$, and $10\downarrow00\uparrow01$), and therefore $2^4=16$ candidate graphs.\\

@@

We now ask which $L=2$ graphs should be used as our new 'base' for the $L=3$ graphs. It is insufficient to use only the $L=2$ graphs that are themselves realizable, as it is not guaranteed that a derived $l=3$ graph will also be impossible.\\
We observe that, when extending a $l=2$ avalanche to a $l=3$ avalanche, most of the conditions required for the $l=2$ avalanche are still included in the conditions for the $l=3$ avalanche (see SI). The only exception is the final set of relaxation inequalities that describes the stability of the state $S^{(2)}$, which is stable for the $l=2$ transition but unstable for the $l=3$ transition. As a result, we can predict that for a $L=2$ graph of which the base inequalities and the first set of relaxation inequalities are incompatible, any derived graphs with longer avalanches must also be impossible.\\
We can thus distinguish three groups of $L=2$ graphs, as indicated in Fig. \ref{fig:N2l2}. The first group of graphs is realizable, and can therefore also be used as a base to construct $L=3$ graphs. For the second group, the graphs themselves are not realizable, but derived $L=3$ graphs may be realizable. For the third group, neither the $L=2$ graphs themselves nor any derived graphs containing longer avalanches are realizable. To construct all possible $l=3$ graphs, we only need to use graphs from the first two groups.\\
We now generalize our method. To find all  $L=l$ graphs, we construct the graphs for $L=l-1$, and divide the graphs into the three groups as described. We then construct all $L=l$ candidate graphs for each of the $l=l-1$ graphs in the first two groups. We can thus iterate over $L$ to find all possible graphs. Our algorithm terminates when none of the graphs belonging to the first and second groups have any possible avalanches of length $l$.

>>>>

The system's response to a field $U$ is thus obtained by comparing the field to the switching fields $\{U_i^+\}, \{U_i^-\}$ of the hysterons individually, flipping hysterons that are unstable, and repeating this process until there are no more unstable hysterons. The response is ill-defined if multiple hysterons are unstable at the same time, or if the system becomes caught in an infinite self-loop.\\
The possible responses of a collection of $N$ hysterons to driving are captured by a state transition graph, which consists of all stable states $S$ and all transitions between states. For a system of $N$ hysterons, there are $2^N$ possible states. In particular there are two saturated states, where either all $s_i=0$ or all $s_i=1$. Each state $S$ has a single up and down transition, except for the two saturated states, which only have an up and a down transition respectively.  Therefore, a transition graph for a collection of $N$ hysterons consists of a maximum of $2^N$ states and $2(2^N-1)$ transitions. It can occur that not all of the $2^N$ states are included in the transition graph, because it is possible for a state to not be reachable, or to only be reachable as an unstable intermediate state of an avalanche. We discuss unreachable states in more detail in Section \ref{ssec:candidate_graphs}.

@@

The region in parameter space associated with each graph is bounded by a set of linear inequalities that the switching fields must obey, which we refer to as the design inequalities. We generate the design inequalities systematically by combining the conditions on each individual transition. We describe how the design inequalities are found in the Supplemental Information. If the solution space of a graph's design inequalities is not empty, then the graph is realizable. Thus, one can find the possible graphs for $N$ hysterons by first generating the candidate graphs, then checking whether each graph's design inequalities have a solution.\\

@@

It often occurs that one or more of the possible states for $N$ hysterons cannot be reached from the ground state. As is customary\cite{juleslechenault2021}, we refer to states that cannot be reached as garden-of-Eden (GoE) states. We may choose to exclude GoE states in some cases, for instance in an experimental setup where one starts from the ground state. As a consequence of excluding GoE states, when two graphs only differ from one another by transitions that occur from GoE states, the graphs become duplicates. Therefore, when GoE states are ignored, the number of distinct graphs can turn out lower than the number of possible combinations of transitions. For $N=2$, ignoring GoE states  reduces the number of graphs from 1458 to 752. 

@@

Using base graphs, we can improve upon our method for generating candidate graphs. Namely, by first constructing all base graphs, and only then evaluating all possible avalanches for each base graph individually, we exclude the subgraphs of the type shown in Fig. \ref{fig:impossiblegraphs}b from the outset. Because the number of graphs for which the design inequalities need to be explicitly generated is then reduced, the base graph method has an improved efficiency compared to the naive approach. \\

@@

In order to discuss scrambling more explicitly, we here define the main loop as the series of up and down transitions connecting the two saturated states. In the Preisach model, the ordering of the main loop fixes the transitions throughout the transition graph, so that each main loop corresponds to a unique Preisach graph. We additionally define the up and down boundary of the main loop as the series of up (down) transitions going from the state with all $s_i=0$ (1) to that with all $s_i=1$ (0). The number of main loops is then the number of permutations of the up boundary times that of the down boundary. Because both the up and down boundary consist of $N$ flips, there are $N!^2$ main loops. However, we fix the up boundary when accounting for exchange symmetry, thus dividing the number of main loops by $N!$.  Therefore, when accounting for exchange symmetry, there are only $N!$ main loops and resulting Preisach graphs.\\
In a scrambled graph, one or more of the transitions in the graph deviates from the main loop, so that each main loop corresponds to a number of base graphs (Fig.\ref{fig:basegraphs_n3}).\\

@@

Let a base graph contain the up transition $S^{(0)}\uparrow S^{(1)}$. From the state $S^{(1)}$ there are two possible transitions: an up transition $S^{(1)}\uparrow S^{(2+)}$, and a down transition  $S^{(1)}\downarrow S^{2-}$.  The transition $S^{(0)}\uparrow S^{(1)}$ can thus be extended to a $l=2$ avalanche in two ways: either by the ferromagnetic avalanche $S^{(0)}\uparrow S^{(1)}\uparrow S^{(2+)}$, or by the antiferromagnetic avalanche $S^{(0)}\uparrow S^{(1)}\downarrow S^{(2-)}$ (Fig. \ref{fig:avalanches}a). Similarly, a down transition $S^{(0)}\downarrow S^{(1)}$ can be extended to the antiferromagnetic avalanche $S^{(0)}\downarrow S^{(1)}\uparrow S^{(2+)}$, or to the ferromagnetic avalanche  $S^{(0)}\downarrow S^{(1)}\downarrow S^{(2-)}$ (Fig. \ref{fig:avalanches}b). Therefore, for each transition $S^{(0)}\rightarrow S^{(1)}$, there is one possible ferromagnetic and one antiferromagnetic avalanche when $S^{(1)}$ is not a saturated state. In the case where $S^{(1)}$ is saturated, only the antiferromagnetic option exists. We must check for each of these transitions that $S^{(2)} \neq S^{(0)}$, because an avalanche cannot contain any cycles. We thus obtain a list of avalanches for each base graph, from which we construct a list of candidate graphs by evaluating each combination of avalanches.\\
An example for one of the $N=2$ base graphs is shown in Fig. \ref{fig:N2l2}. For the example shown, the base graph has four possible $l=2$ avalanches all associated with a different transition ($00\uparrow 01\uparrow 11$, $01\uparrow11\downarrow10$, $11\downarrow10\downarrow00$, and $10\downarrow00\uparrow01$), and therefore $2^4=16$ candidate graphs.\\

@@

We now ask which $L=2$ graphs should be used as our new 'base' for the $L=3$ graphs. It is insufficient to use only the $L=2$ graphs that are themselves realizable, as it is not guaranteed that a derived $L=3$ graph will also be impossible.\\
We observe that, when extending a $l=2$ avalanche to a $l=3$ avalanche, most of the conditions required for the $l=2$ avalanche are still included in the conditions for the $l=3$ avalanche (see SI). The only exception is the final set of relaxation inequalities that describes the stability of the state $S^{(2)}$, which is stable for the $l=2$ transition but unstable for the $l=3$ transition. As a result, we can predict that for a $L=2$ graph of which the base inequalities and the first set of relaxation inequalities are incompatible, any derived graphs with longer avalanches must also be impossible. If, however, it is only the last set of relaxation inequalities that causes the graph to be impossible, then a derived graph of $L=3$ or higher may still be realizable.\\
We can thus distinguish three groups of $L=2$ graphs, as indicated in Fig. \ref{fig:N2l2}. The first group of graphs is realizable, and can therefore also be used as a base to construct $L=3$ graphs. For the second group, the graphs themselves are not realizable, but derived $L=3$ graphs may be realizable. For the third group, neither the $L=2$ graphs themselves nor any derived graphs containing longer avalanches are realizable. To construct all possible $L=3$ graphs, we only need to use graphs from the first two groups.\\
We now generalize our method. To find all  $L=l$ graphs, we construct the graphs for $L=l-1$, and divide the graphs into the three groups as described. We then construct all $L=l$ candidate graphs for each of the $l=l-1$ graphs in the first two groups. We thus have an algorithm that constructs all realizable graphs by iterating over the maximum avalanche length $L$. Our algorithm terminates when none of the graphs belonging to the first and second groups have any possible avalanches of length $l$.
